---
title: "Crime in Newfoundland and Labrador (2006-2020)"
output:
  html_document:
    df_print: paged
---

```{r}
  library(tidyverse)
  library(janitor)
  library(ggthemes)

  data <- read.csv("https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=17253")

  data <- data|>
    janitor::clean_names() |>
    mutate(geography=factor(geography)) |>
    # Focus upon actual incidents of crime.
    filter(data_type=="actual_incidents") |>
    # We will restrict our area of study to the police regions of Newfoundland
    # and Labrador.
    filter(
      geography != "Yukon",
      geography != "Saskatchewan",
      geography != "Quebec",
      geography != "Prince Edward Island",
      geography != "Ontario",
      geography != "Nunavut",
      geography != "Northwest Territories",
      geography != "New Brunswick",
      geography != "Manitoba",
      geography != "Canada",
      geography != "British Columbia",
      geography != "Alberta",
      geography != "Nova Scotia",
      geography != "Newfoundland and Labrador",
    ) |>
    select(!c('geography_type','geography_note','data_type')) |>
    mutate(geography = str_replace(geography,"Northeast Avalon R.N.C","Northeast Avalon, R.N.C."))
```

This [RStudio](https://resources.rstudio.com/categories/rstudio-ide/) notebook studies crime in Newfoundland and Labrador over a 14 year period, using a data set made available by **Open Data Newfoundland and Labrador** on their website.

# Violent Crime

## Which region experienced the largest number of homicides between 2006-2020?

```{r}
  data |>
    ggplot(aes(reorder(geography,homicide), homicide)) +
    geom_col(aes(fill="Homicides",color="Homicides")) +
    labs(
      title="Homicides by Region",
      x="Region",
      y="Total Homicides",
      subtitle="(2006-2020)",
      caption="Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    theme(legend.position = "none") +
    scale_color_colorblind() +
    scale_fill_colorblind()
```

The **Northeast Avalon** region experienced the largest number of homicides over a 14-year period.

## How did the total number of homicides in the Northeast Avalon region vary over time?

```{r}
  data |>
  filter(geography=="Northeast Avalon, R.N.C.") |>
  ggplot(aes(year,homicide)) +
  geom_line() +
  geom_smooth(se=FALSE,method="lm") +
  labs(
    title="Total Homicides in Northeast Avalon Region Over Time",
    x="Year",
    y="Total Homicides",
    caption="Data from Open Data Newfoundland and Labrador",
  )
```

There is a **positive** trend in the number of homicides in the Northeast Avalon region.

## Is this a sign of an increase in violent crimes in general for the Northeast Avalon region?

We can get a sense for the amount of overall violent crime in the region by counting the number of total violent criminal code violations over time.

```{r}
  data |>
  filter(geography=="Northeast Avalon, R.N.C.") |>
  ggplot(aes(year,total_violent_criminal_code_violations)) +
  geom_line() +
  geom_smooth(se=FALSE,method="lm") +
  labs(
    title="Total Violent Crimes in Northeast Avalon Region Over Time",
    x="Year",
    y="Total Violent Criminal Code Violations",
    caption="Data from Open Data Newfoundland and Labrador",
  )
```

There is a **positive** trend in the number of total violent criminal code violations for the region. This does seem to support the notion that violent crime has, in general, increased for the Northeast Avalon region over time.

## How do the various regions rank in terms of total violent criminal code violations?

```{r}
data |>
  ggplot(aes(reorder(geography,total_violent_criminal_code_violations),total_violent_criminal_code_violations)) +
  geom_col(aes(color="Violations",fill="Violations")) +
  labs(
    title="Total Violent Criminal Code Violations by Region",
    x="Region",
    y="Total Violent Criminal Code Violations",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  coord_flip()
```

The **Northeast Avalon** region leads in terms of total violent criminal code violations.

## Which region leads in assaults against peace officers?

```{r}
data |>
  ggplot(aes(reorder(geography,assault_peace_officer),assault_peace_officer)) +
  geom_col(aes(color="Assaults",fill="Assaults")) +
  coord_flip() +
  labs(
    title="Total Assaults Against Peace Officers by Region",
    x="Region",
    y="Total Assaults Against Peace Officers",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

Here we see that the **Northeast Avalon** region leads in terms of assaults against peace officers. This certainly fits with the general trend of increasing violent crime within the region. It is concerning, however, to note the number of incidents in which even officers of the law became the victims of violent crimes.

# Sex Crimes

## How do the regions break down in terms of sexual assaults?

Sexual assaults are "ranked" in three levels:

- level 1,
- level 2 (committed with a weapon or causing bodily harm),
- level 3 (aggravated).

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_1),sexual_assault_level_1)) +
    geom_col(aes(fill="Assaults",color="Assaults")) +
    labs(
      title="Sexual Assaults (Level 1) by Region",
      x = "Sexual Assaults (Level 1)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Northeast Avalon** region ranks first in terms of total number of sexual assaults (level 1).

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_2_with_a_weapon_or_bodily_harm),sexual_assault_level_2_with_a_weapon_or_bodily_harm)) +
    geom_col(aes(fill="Assaults",color="Assaults"))+
    labs(
      title="Sexual Assaults (Level 2) by Region",
      x = "Sexual Assaults (Level 2)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

Again, the **Northeast Avalon** leads in terms of sexual assaults, this time involving a weapon or bodily harm.

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_3_aggravated),sexual_assault_level_3_aggravated)) +
    geom_col(aes(fill="Assaults",color="Assaults")) +
    labs(
      title="Sexual Assaults (Level 3) by Region",
      x = "Sexual Assaults (Level 3)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Labrador District** experienced the greatest number of aggravated sexual assaults (level 3).

## How are children affected by sex crimes in Newfoundland and Labrador?

```{r}
  data |>
    ggplot(aes(reorder(geography,total_sexual_violations_against_children),total_sexual_violations_against_children)) +
    geom_col(aes(fill="Violations",color="Violations")) +
    labs(
      title="Total Sexual Violations Against Children by Region",
      x = "Sexual Violations",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Northeast Avalon** district experienced the most total sexual violations against children.

In addition, there are incidences of child luring via computer to be considered.

```{r}
  data |>
    ggplot(aes(reorder(geography,luring_a_child_via_a_computer),luring_a_child_via_a_computer)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    labs(
      title="Total Incidents of Child Luring via Computer by Region",
      x = "Incidents",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

Here it is apparent that the **Northeast Avalon** region leads by a wide margin in terms of child luring incidents via computer.

## Is there a trend in sexual violations against children in the Northeast Avalon region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_sexual_violations_against_children)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Sexual Violations Against Children",
      subtitle="Northeast Avalon Region",
      x="Year",
      y="Sexual Violations",
      caption="Data from Open Data Newfoundland and Labrador",
    )
```

There is a clear, **positive** trend in the number of sexual violations against children, with respect to the Northeast Avalon region.

