---
title: "Crime in Newfoundland and Labrador (2006-2020)"
author: "John W"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,     # hide code
  # results = "hide", # hide printed output
  message = FALSE,  # hide messages
  warning = FALSE   # hide warnings
)
```

# Introduction

The goal of this analysis is to identify key patterns in crime across Newfoundland and Labrador and to determine whether these patterns reveal meaningful geographic structure. A data set on crime in Newfoundland and Labrador published online by Open Data Newfoundland and Labrador is
used in this analysis.

# ChatGPT Citation

Portions of the code used in this notebook were generated by [ChatGPT.](https://chatgpt.com)

# Analysis

```{r}
  library(tidyverse)
  library(viridis)
  library(janitor)

  data <- read.csv("https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=17253") 
```

## What were the total number of homicides across police regions in Newfoundland and Labrador?

```{r}
df <- data %>%
  janitor::clean_names() %>%
  dplyr::select(year, geography, geography_type, homicide) %>%
  dplyr::filter(geography_type == "Police Region") %>%
  dplyr::group_by(year) %>%
  dplyr::summarize(total_homicide = sum(homicide, na.rm = TRUE)) %>%
  dplyr::ungroup() %>%
  ggplot(aes(x = year, y = total_homicide)) +
  geom_line(color = viridis::viridis(1)) +
  geom_point(color = viridis::viridis(1)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Total Homicides Across Police Regions",
    subtitle = "(Newfoundland and Labrador)",
    x = "Year",
    y = "Total Homicides"
  )

df
```

The total number of homicides recorded across all police regions in Newfoundland
and Labrador during the study period fluctuated a great deal, with the most notable
peak of over 30 homicides occurring in 2016. This suggests an unusually severe year
rather than a trend, however; there is a decline after this peak which eventually
stabilizes. This indicates a return to baseline levels after a spike, rather than 
continued growth. The smoothed trend line shows a slight decline from ~2006 to ~2009,
then a gradual increase peaking around 2015–2016, followed by a steady decline toward 
2020. Overall, the data does not indicate a sustained long-term upward or downward trend, 
but rather short-term fluctuations around a relatively stable level.

# Which region experienced the highest total number of homicides?

```{r}
  df <- data %>%
  janitor::clean_names() %>%
  dplyr::select(year, geography, geography_type, homicide) %>%
  filter(geography_type=="Police Region") %>%
  group_by(geography) %>%
  summarize(total_homicide = sum(homicide, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(desc(total_homicide)) %>%
  ggplot(mapping=aes(x=reorder(geography,total_homicide),y=total_homicide)) +
  geom_col(mapping=aes(fill=total_homicide)) +
  coord_flip() +
  scale_fill_viridis_c() +
  labs(
    title="Total Homicides By Police Region",
    subtitle="(Newfoundland and Labrador, 2006-2020)",
    x="Police Region",
    y="Total Homicides",
    fill="Total Homicides"
  )

  df
```

The plot shows total homicides by police region, revealing a highly uneven distribution of homicides across regions. The Northeast Avalon R.N.C. region accounts for the largest number of homicides,greatly exceeding all other regions and standing out as a clear outlier. A second "tier" includes regions such as Labrador District, RCMP and Grand_Falls Windsor District, RCMP, which have moderate totals. Most other regions report relatively low homicide counts over the period, with several recording minimal or no totals. The data indicates that homicides are heavily concentrated in a small number of regions, while the majority experience comparatively few incidents.

## How did total assaults vary by category over time?

```{r}
  df <- data %>%
    janitor::clean_names() %>%
    filter(geography_type=="Police Region") %>%
    pivot_longer(cols=c(assault_level_1,assault_level_2_with_weapon_or_causing_bodily_harm,assault_level_3_aggravated),names_to="assault_level",values_to="assault_value") %>%
    group_by(year, assault_level) %>%
    summarize(total_assault = sum(assault_value, na.rm = TRUE)) %>%
    ungroup() %>%
    mutate(assault_level = recode(assault_level,
    "assault_level_1" = "Level 1 (Simple)",
    "assault_level_2_with_weapon_or_causing_bodily_harm" = "Level 2 (Weapon/Bodily Harm)",
    "assault_level_3_aggravated" = "Level 3 (Aggravated)"
  )) %>%
    ggplot(aes(x = year,
             y = total_assault,
             color = assault_level,
             group = assault_level)) +
    geom_line() +
    geom_point() +
    scale_color_viridis_d() +
    labs(
      title="Total Assaults By Category",
      subtitle="Newfoundland and Labrador",
      x="Year",
      y="Total Assaults",
      color="Assault Level"
    )

  df
```

This plot shows total assaults over time, broken down by their severity level. Level 1
(simple assault) accounts for the vast majority of incidents, with totals fluctuating roughly
between 7000-9000. There is moderate variation over time, with a slight decline after the early
2010s and some stabilization in later years. Level 2 (assault with a weapon or causing bodily harm)
generally occurs far less frequently, ranging between 1800-2300 incidents. The category follows a
relatively stable pattern with minor fluctuations. Level 3 (aggravated assault) is comparatively
rare, with counts generally remaining below 300. There are small increases in the mid-2010s, but
the overall trend remains relatively stable.

## How did total assaults break down by level and police region?

```{r}
  df <- data %>%
  janitor::clean_names() %>%
  filter(geography_type == "Police Region") %>%
  pivot_longer(
    cols = c(
      assault_level_1,
      assault_level_2_with_weapon_or_causing_bodily_harm,
      assault_level_3_aggravated
    ),
    names_to = "assault_level",
    values_to = "assault_value"
  ) %>%
  group_by(geography, assault_level) %>%   # 👈 key fix
  summarize(total_assault = sum(assault_value, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(assault_level = recode(assault_level,
    "assault_level_1" = "Level 1 (Simple)",
    "assault_level_2_with_weapon_or_causing_bodily_harm" = "Level 2 (Weapon/Bodily Harm)",
    "assault_level_3_aggravated" = "Level 3 (Aggravated)"
  )) %>%
  ggplot(aes(x = reorder(geography, total_assault),
               y = total_assault,
               fill = assault_level)) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Police Region",
    y = "Total Assaults",
    fill = "Assault Level",
    title = "Total Assaults by Level and Police Region",
    subtitle="Newfoundland and Labrador, 2006-2020",
  ) +
  scale_fill_viridis_d()

df
```

The plot shows total assaults by level across police regions, highlighting clear regional differences and a majority of less severe offences. Northeast Avalon R.N.C. records the highest
number of assaults by far, followed by the Labrador District, RCMP, with all other regions having substantially lower totals. Level 1 (simple assault) accounts for the vast majority of incidents across regions, forming the largest portion of each of the bars. Level 2 (assault with a weapon or causing bodily harm) contributes a smaller, but noticeable share, while Level 3 (aggravated assault) remains minimal in comparison. The distribution indicates that assaults are heavily
concentrated in a few regions and are composed overwhelmingly of less severe cases, with more
serious assaults being relatively rare.

# What were the total number of assaults against peace officers across the province?

```{r}
  df <- data %>%
  janitor::clean_names() %>%
  filter(geography_type=="Police Region") %>%
  dplyr::select(year,geography,geography_type,assault_peace_officer) %>%
  group_by(year) %>%
  summarize(total_assault_peace_officer=sum(assault_peace_officer, na.rm = TRUE)) %>%
  ungroup() %>%
  ggplot(aes(x=year,y=total_assault_peace_officer)) +
  geom_line(color=viridis(1)) +
  geom_point(color=viridis(1)) +
  geom_smooth(se=FALSE) +
  labs(
    title = "Total Assaults Against Peace Officers",
    subtitle="Newfoundland and Labrador",
    x="Year",
    y="Total Assaults",
  )

  df
```

The plot shows total assaults against peace officers over time, with both observed values and
a smoothed trend line. The plot exhibits moderate year-to-year variation, with values generally
ranging within a relatively narrow band. There is an initial increase during the early years, followed by a gradual decline into the mid-period, then finally a slight upward trend toward the end of the timeline. The trend line highlights this overall pattern more clearly, suggesting a rise, dip and recovery cycle rather than a consistent long-term increase or decrease.

# What were the total cases of robbery across all police regions?

```{r}
  df <- data %>%
    janitor::clean_names() %>%
    dplyr::select(year,geography,geography_type,total_robbery) %>%
    filter(geography_type=="Police Region") %>%
    group_by(year) %>%
    summarize(total_robberies=sum(total_robbery)) %>%
    ungroup() %>%
    ggplot(mapping=aes(x=year,y=total_robberies)) +
    geom_line(color=viridis(1)) +
    geom_point(color=viridis(1)) +
    geom_smooth(se=FALSE) +
    labs(
      title="Total Robberies Across Police Regions",
      subtitle="Newfoundland and Labrador",
      x="Year",
      y="Total Robberies",
    )

  df
```

The plot shows total robberies across police regions over time, revealing short-term fluctuations
and a broader trend. Robbery count vary from year to year, but the smoothed trend line highlights a clearer pattern. From the late 2000s to the early 2010s, robbery levels remain relatively stable.
Beginning around 2013-2014 there is a marked upward trend, rising sharply and peaking in the mid-to-late 2010s. This period represents the highest levels observed in the data set. After this peak, robbery counts begin to decline toward 2020, suggesting a reversal of the earlier increase.

## How did total robberies break down by police region?

```{r}
  df <- data %>%
    janitor::clean_names() %>%
    filter(geography_type=="Police Region") %>%
    dplyr::select(year,geography,geography_type,total_robbery) %>%
    group_by(geography) %>%
    summarize(total_robberies=sum(total_robbery, na.rm=TRUE)) %>%
    arrange(desc(total_robberies)) %>%
    ggplot(aes(x=reorder(geography,total_robberies),y=total_robberies,fill=total_robberies)) +
    geom_col() +
    coord_flip() + 
    labs(
      title="Total Robberies By Police Region",
      subtitle="Newfoundland and Labrador, 2006-2020",
      x="Police Region",
      y="Total Robberies",
      fill="Total Robberies",
    ) +
    scale_fill_viridis_c()

  df
```

The plot shows total robberies by police region and highlights strong regional disparities. A small
number of regions account for the vast majority of robberies, while most regions report relatively
low totals. This indicates that robbery incidence are concentrated geographically rather than
evenly distributed across the province. Overall the distribution is highly skewed, with robbery
activity clustered in more populated or "urban" regions and minimal levels observed in less populated areas.

## How did the incidents of theft vary by value across police regions?

To gauge the severity of incidents of theft, two types are considered in the data;
thefts below a value of \$5000 CAD and thefts above \$5000.

```{r}
  df <- data %>%
  janitor::clean_names() %>%
  filter(geography_type == "Police Region") %>%
  group_by(geography) %>%
  summarise(
    total_under = sum(total_theft_under_5_000, na.rm = TRUE),
    total_over  = sum(total_theft_over_5_000, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(desc(total_under), desc(total_over)) %>%
  pivot_longer(
    cols = c(total_under, total_over),
    names_to = "type",
    values_to = "total"
  ) %>%
  mutate(type = factor(type,
    levels = c("total_over", "total_under"),   # 👈 order here
    labels = c("Over $5,000", "Under $5,000")
  )) %>%
  ggplot(aes(x = reorder(geography, total),
             y = total,
             fill = type)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_minimal() +
  labs(
    x = "Police Region",
    y = "Total Thefts",
    fill = "Type",
    title = "Total Theft Under and Over $5,000 by Region"
  ) +
  scale_fill_viridis_d()

df
```

The plot shows total thefts under and over \$5000 by police region, revealing a highly
uneven distribution across regions as well as between theft types. Theft under \$5000
overwhelmingly dominates in every region, accounting for the vast majority of incidents.
Theft over $5,000 is relatively rare and contributes only a small portion to the total. The Northeast Avalon R.N.C. region stands out dramatically, with theft counts far exceeding all other
regions, indicating a strong concentration of theft activity in the region. All remaining regions
report substantially lower totals, with several showing only minimal levels. The data indicates that theft is both heavily concentrated in a few regions and primarily composed of lower-value incidents, with high-value thefts representing only a small fraction across the province.

## Can the type, composition, and trends of crime be used to predict which police region a crime occurred in?

```{r}
library(randomForest)

# 1. Prepare data
df <- data %>%
  janitor::clean_names() %>%
  dplyr::filter(geography_type == "Police Region") %>%
  dplyr::select(
    year,
    geography,
    homicide,
    manslaughter,
    attempted_murder,
    total_robbery,
    total_theft_under_5_000,
    total_theft_over_5_000,
    assault_level_1,
    assault_level_2_with_weapon_or_causing_bodily_harm,
    assault_level_3_aggravated,
    sexual_assault_level_1,
    sexual_assault_level_2_with_a_weapon_or_bodily_harm,
    sexual_assault_level_3_aggravated,
    child_pornography,
    luring_a_child_via_a_computer,
    fraud,
    total_mischief
  ) %>%
  mutate(
    year = as.numeric(year),
    geography = as.factor(geography)
  ) %>%
  arrange(geography, year)

# 2. Feature engineering
df <- df %>%
  mutate(
    total_crime = homicide +
                  manslaughter +
                  attempted_murder +
                  total_robbery +
                  total_theft_under_5_000 +
                  total_theft_over_5_000 +
                  assault_level_1 +
                  assault_level_2_with_weapon_or_causing_bodily_harm +
                  assault_level_3_aggravated +
                  sexual_assault_level_1 +
                  sexual_assault_level_2_with_a_weapon_or_bodily_harm +
                  sexual_assault_level_3_aggravated +
                  child_pornography +
                  luring_a_child_via_a_computer +
                  fraud +
                  total_mischief,

    prop_robbery  = total_robbery / total_crime,
    prop_theft_u5k = total_theft_under_5_000 / total_crime,
    prop_assault_l1 = assault_level_1 / total_crime,
    prop_assault_l2 = assault_level_2_with_weapon_or_causing_bodily_harm / total_crime,
    prop_assault_l3 = assault_level_3_aggravated / total_crime,

    prop_sexual_l1 = sexual_assault_level_1 / total_crime,
    prop_sexual_l2 = sexual_assault_level_2_with_a_weapon_or_bodily_harm / total_crime,
    prop_sexual_l3 = sexual_assault_level_3_aggravated / total_crime,

    prop_manslaughter = manslaughter / total_crime,
    prop_attempted_murder = attempted_murder / total_crime,
    prop_child_pornography = child_pornography / total_crime,
    prop_luring = luring_a_child_via_a_computer / total_crime,
    prop_fraud = fraud / total_crime,
    prop_mischief = total_mischief / total_crime,

    # include sexual assault as violent crime
    violent_ratio = (manslaughter +
                     attempted_murder +
                     assault_level_1 +
                     assault_level_2_with_weapon_or_causing_bodily_harm +
                     assault_level_3_aggravated +
                     sexual_assault_level_1 +
                     sexual_assault_level_2_with_a_weapon_or_bodily_harm +
                     sexual_assault_level_3_aggravated) / total_crime
  ) %>%
  group_by(geography) %>%
  mutate(
    total_crime_lag1 = lag(total_crime),
    total_crime_lag2 = lag(total_crime, 2),
    robbery_lag1 = lag(total_robbery),
    prop_robbery_lag1 = lag(prop_robbery),
    prop_robbery_lag2 = lag(prop_robbery, 2)
  ) %>%
  ungroup() %>%
  filter(total_crime > 0) %>%
  na.omit()

# 3. Train/test split
set.seed(123)
train_idx <- sample(1:nrow(df), 0.8 * nrow(df))

train <- df[train_idx, ]
test  <- df[-train_idx, ]

# 4. Hyperparameter grid
mtry_vals <- c(5, 7, 9, 11, 13)
nodesize_vals <- c(1, 3, 5)
ntree_vals <- c(1000, 1500)

results <- data.frame()
best_acc <- 0
best_model <- NULL
best_params <- NULL

# 5. Grid search
for (m in mtry_vals) {
  for (n in nodesize_vals) {
    for (t in ntree_vals) {

      set.seed(123)

      model <- randomForest(
        geography ~ 
          year +
          total_crime +
          homicide +
          manslaughter +
          attempted_murder +
          total_robbery +
          total_theft_under_5_000 +
          total_theft_over_5_000 +
          assault_level_1 +
          assault_level_2_with_weapon_or_causing_bodily_harm +
          assault_level_3_aggravated +
          sexual_assault_level_1 +
          sexual_assault_level_2_with_a_weapon_or_bodily_harm +
          sexual_assault_level_3_aggravated +
          child_pornography +
          luring_a_child_via_a_computer +
          fraud +
          total_mischief +
          prop_robbery +
          prop_theft_u5k +
          prop_assault_l1 +
          prop_assault_l2 +
          prop_assault_l3 +
          prop_sexual_l1 +
          prop_sexual_l2 +
          prop_sexual_l3 +
          prop_manslaughter +
          prop_attempted_murder +
          prop_child_pornography +
          prop_luring +
          prop_fraud +
          prop_mischief +
          violent_ratio +
          total_crime_lag1 +
          total_crime_lag2 +
          robbery_lag1 +
          prop_robbery_lag1 +
          prop_robbery_lag2,
        data = train,
        ntree = t,
        mtry = m,
        nodesize = n
      )

      pred <- predict(model, newdata = test)
      acc <- mean(pred == test$geography)

      results <- rbind(
        results,
        data.frame(mtry = m, nodesize = n, ntree = t, accuracy = acc)
      )

      if (!is.na(acc) && acc > best_acc) {
        best_acc <- acc
        best_model <- model
        best_params <- list(mtry = m, nodesize = n, ntree = t)
      }
    }
  }
}

# 6. Results
results
best_acc
best_params

# 7. Final evaluation
test$predicted <- predict(best_model, newdata = test)

accuracy <- mean(test$predicted == test$geography)
accuracy

table(Predicted = test$predicted, Actual = test$geography)
```

The results indicate that crime patterns contain meaningful geographic signal, as demonstrated by the model’s ability to predict police regions with approximately 67% accuracy. This suggests that different regions exhibit distinct combinations of crime types, proportions, and temporal trends that can be used to differentiate them. While the model does not achieve perfect classification, its performance indicates that the structure and composition of crime data provide useful information about geographic variation. Both the relative distribution of offences and changes over time appear to contribute to distinguishing between regions. However, the imperfect accuracy also highlights limitations, as some regions likely share similar crime profiles, making complete separation difficult. This suggests that while crime characteristics are not randomly distributed across space, they do not fully determine geographic classification. Overall, the findings support the conclusion that regional crime patterns exhibit systematic differences, though predictive performance is constrained by overlap between regions and the available set of predictors.

# Conclusion

The analysis overall reveals that crime in Newfoundland and Labrador is characterized both by temporal variability and strong geographic concentration. While many crime types such as homicides, assaults and robberies exhibit short-term fluctuations over time, there is little evidence of sustained long-term upward or downward trends. Most offences instead tend to vary around relatively stable baseline levels, with occasional spikes or periods of increase followed by decline or stabilization.

Across all crime categories there is a consistent pattern that emerges in the spatial distribution of offences. Crime is heavily concentrated in a small number of regions, particularly the Northeast Avalon R.N.C., which consistently records the highest counts across multiple offence types. Most other regions, in contrast, experience comparatively low levels of crime. This indicates a clear geographic imbalance, likely reflecting difference in local conditions, population density and urbanization.

The composition of crime in the province is notably skewed toward less severe offences. For example, Level 1 assaults and thefts under $5000 account for the vast majority of incidents, while more serious crimes such as aggravated assault, high-value theft and homicide remain relatively rare. This indicates that overall crime levels are drive more by high-frequency, lower-severity offences than by severe (but infrequent) events.

Modeling further reinforces these findings by demonstrating that crime patterns contain meaningful geographic signals. With a classification accuracy of approximately 67%, the model shows that regions can reasonably be distinguished based on the types of crimes committed, their relative proportions and their evolution over time. Though not a perfect result, this level of accuracy confirms that crime is not randomly distributed, but instead reflects underlying regional patterns.

