---
title: "Crime in Newfoundland and Labrador (2006-2020)"
output:
  html_document:
    df_print: paged
---

```{r}
  library(tidyverse)
  library(janitor)
  library(ggthemes)

  data <- read.csv("https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=17253")

  data <- data|>
    janitor::clean_names() |>
    mutate(geography=factor(geography)) |>
    # Focus upon actual incidents of crime.
    filter(data_type=="actual_incidents") |>
    # We will restrict our area of study to the police regions of Newfoundland
    # and Labrador.
    filter(
      geography != "Yukon",
      geography != "Saskatchewan",
      geography != "Quebec",
      geography != "Prince Edward Island",
      geography != "Ontario",
      geography != "Nunavut",
      geography != "Northwest Territories",
      geography != "New Brunswick",
      geography != "Manitoba",
      geography != "Canada",
      geography != "British Columbia",
      geography != "Alberta",
      geography != "Nova Scotia",
      geography != "Newfoundland and Labrador",
    ) |>
    select(!c('geography_type','geography_note','data_type')) |>
    mutate(geography = str_replace(geography,"Northeast Avalon R.N.C","Northeast Avalon, R.N.C."))
```

This [RStudio](https://resources.rstudio.com/categories/rstudio-ide/) notebook studies crime in Newfoundland and Labrador over a 14 year period, using a data set made available by **Open Data Newfoundland and Labrador** on their website.

# Violent Crime

## Which region experienced the largest number of homicides between 2006-2020?

```{r}
  data |>
    ggplot(aes(reorder(geography,homicide), homicide)) +
    geom_col(aes(fill="Homicides",color="Homicides")) +
    labs(
      title="Homicides by Region",
      x="Region",
      y="Total Homicides",
      subtitle="(2006-2020)",
      caption="Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    theme(legend.position = "none") +
    scale_color_colorblind() +
    scale_fill_colorblind()
```

The **Northeast Avalon** region experienced the largest number of homicides over a 14-year period.

## How did the total number of homicides in the Northeast Avalon region vary over time?

```{r}
  data |>
  filter(geography=="Northeast Avalon, R.N.C.") |>
  ggplot(aes(year,homicide)) +
  geom_line() +
  geom_smooth(se=FALSE,method="lm") +
  labs(
    title="Total Homicides in Northeast Avalon Over Time",
    x="Year",
    y="Total Homicides",
    caption="Data from Open Data Newfoundland and Labrador",
  ) + 
  scale_x_continuous(breaks=seq(2006,2020,1))
```

There is a **positive** trend in the number of homicides in the **Northeast Avalon** region.

## Which region leads in assaults against peace officers?

```{r}
data |>
  ggplot(aes(reorder(geography,assault_peace_officer),assault_peace_officer)) +
  geom_col(aes(color="Assaults",fill="Assaults")) +
  coord_flip() +
  labs(
    title="Total Assaults Against Peace Officers by Region",
    x="Region",
    y="Total Assaults Against Peace Officers",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

Here we see that the **Northeast Avalon** region leads in terms of assaults against peace officers. This certainly fits with the general trend of increasing violent crime within the region. It is concerning, however, to note the number of incidents in which even officers of the law became the victims of violent crimes.

## Is there a trend in the number of assaults against peace officers for the Northeast Avalon region?

```{r}
  data |>
  filter(geography=="Northeast Avalon, R.N.C.") |>
  ggplot(aes(year,assault_peace_officer)) +
  geom_line() +
  geom_smooth(se=FALSE,method="lm") +
  labs(
    title="Total Assaults Against Peace Officers in Northeast Avalon Over Time",
    x="Year",
    y="Total Assaults",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  scale_x_continuous(breaks=seq(2006,2020,1))
```

There is a weak, **positive** trend in assaults against peace officers in the **Northeast Avalon** region. This appearance may be deceptive, however, particularly in the case of the "peak" during 2010 when assaults against peace officers were apparently the highest. The [metadata](https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=41) for the data set used in this study notes that, "new legislation in 2009 regarding assaults on peace officers may be the cause of an increase in reported offenses." 

## How many assaults, and of what kind, were committed in the Northeast Avalon region?

Assaults are ranked on a scale. There are three levels:

- level 1,
- level 2 (involving a weapon or causing bodily harm),
- level 3 (aggravated).

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year)) +
    geom_line(aes(y=assault_level_1,color="Assault (Level 1)")) +
    geom_line(aes(y=assault_level_2_with_weapon_or_causing_bodily_harm,color="Assault (Level 2)")) +
    geom_line(aes(y=assault_level_3_aggravated,color="Assault (Level 3)")) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  theme(legend.position="bottom") +
  labs(
    title="Total Assaults Against Members of the Public (Level 1-3)",
    subtitle="Northeast Avalon",
    x="Year",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
    color="Legend"
  ) +
  scale_y_log10() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

Level 1 assaults account for more than any other level in the **Northeast Avalon** region.

# Sex Crimes

## How do the regions break down in terms of sexual assaults?

Sexual assaults are "ranked" by three levels:

- level 1,
- level 2 (committed with a weapon or causing bodily harm),
- level 3 (aggravated).

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_1),sexual_assault_level_1)) +
    geom_col(aes(fill="Assaults",color="Assaults")) +
    labs(
      title="Sexual Assaults (Level 1) by Region",
      x = "Sexual Assaults (Level 1)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Northeast Avalon** region ranks first in terms of total number of sexual assaults (level 1).

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_2_with_a_weapon_or_bodily_harm),sexual_assault_level_2_with_a_weapon_or_bodily_harm)) +
    geom_col(aes(fill="Assaults",color="Assaults"))+
    labs(
      title="Sexual Assaults (Level 2) by Region",
      x = "Sexual Assaults (Level 2)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

Again, the **Northeast Avalon** leads in terms of sexual assaults (level 2), this time involving a weapon or bodily harm.

```{r}
  data |>
    ggplot(aes(reorder(geography,sexual_assault_level_3_aggravated),sexual_assault_level_3_aggravated)) +
    geom_col(aes(fill="Assaults",color="Assaults")) +
    labs(
      title="Sexual Assaults (Level 3) by Region",
      x = "Sexual Assaults (Level 3)",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Labrador District** experienced the greatest number of aggravated sexual assaults (level 3).

## How are children affected by sex crimes in Newfoundland and Labrador?

```{r}
  data |>
    ggplot(aes(reorder(geography,total_sexual_violations_against_children),total_sexual_violations_against_children)) +
    geom_col(aes(fill="Violations",color="Violations")) +
    labs(
      title="Total Sexual Violations Against Children by Region",
      x = "Sexual Violations",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Northeast Avalon** district experienced the most total sexual violations against children.

## What types and amounts of sex crimes against children were experienced in the Northeast Avalon region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year)) +
    geom_line(aes(y=child_pornography,color="Child Pornography")) +
    geom_line(aes(y=luring_a_child_via_a_computer,color="Luring a Child via a Computer")) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  theme(legend.position="bottom") +
  labs(
    title="Sex Crimes Against Children",
    subtitle="Northeast Avalon",
    x="Year",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
    color="Legend"
  ) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

These numbers do not account for the total sex crimes committed against children in the region, only for two specific examples.

## Is there a trend in sexual violations against children in the Northeast Avalon region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_sexual_violations_against_children)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Sexual Violations Against Children",
      subtitle="Northeast Avalon",
      x="Year",
      y="Sexual Violations",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    scale_x_continuous(breaks=seq(2006,2020,1))
```

There is a clear, **positive** trend in the number of sexual violations against children, with respect to the **Northeast Avalon** region.

## How do the charges of prostitution vary by region?

```{r}
  data |>
    ggplot(aes(reorder(geography,total_prostitution),total_prostitution)) +
    geom_col(aes(fill="Charges",color="Charges")) +
    labs(
      title="Total Charges of Prostitution by Region",
      x = "Charges",
      y = "Region",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

In this case the **Northeast Avalon** region outnumbers all others in terms of charges laid for prostitution.

# Robbery and Theft

## How did the number of robberies rank across the different regions over time?

```{r}
  data |>
    ggplot(aes(reorder(geography,total_robbery),total_robbery)) +
    geom_col(aes(fill="Robberies",color="Robberies")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Total Robberies by Region",
      x="Region",
      y="Robberies",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

We see in this case that the **Northeast Avalon** experienced the most robberies over time; in fact, the region outstrips the others by a wide margin in terms of this form of crime.

## Is there a trend in the number of robberies in the Northeast Avalon region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_robbery)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Robberies",
      subtitle="Northeast Avalon",
      x="Year",
      y="Robberies",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    scale_x_continuous(breaks=seq(2006,2020,1))
```

There has been a clear, **positive** trend in the number of robberies over time in the **Northeast Avalon** region. However, it would take more data and further study to determine the exact reasons for such a disparity between this region and the others.

## How did the various regions break down in terms of shoplifting incidents?

```{r}
  data |>
    ggplot(aes(reorder(geography,shoplifting_5_000_or_under),shoplifting_5_000_or_under)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Shoplifting Incidents (Under $5,000)",
      x="Region",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

Here again, we see that the **Northeast Avalon** dominates by a wide margin in terms of the number of total minor shoplifting incidents over time. Is this part of a trend as well?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,shoplifting_5_000_or_under)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Shoplifting Incidents ($5,000 or under)",
      subtitle="Northeast Avalon",
      x="Year",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    scale_x_continuous(breaks=seq(2006,2020,1))
```

The number of minor shoplifting incidents in the **Northeast Avalon** region seem to be trending **downward** over time.

What about the number of serious shoplifting incidents, in which property of over $5,000 in value was stolen?

```{r}
  data |>
    ggplot(aes(reorder(geography,shoplifting_over_5_000),shoplifting_over_5_000)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Shoplifting Incidents (Over $5,000)",
      x="Region",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

Again, we see that the **Northeast Avalon** region vastly outnumbers its neighbors in terms of shoplifting incidents, even when those incidents involve much more expensive property. The trend here may also be revealing.

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,shoplifting_over_5_000)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Shoplifting Incidents (Over $5,000)",
      subtitle="Northeast Avalon",
      x="Year",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    scale_x_continuous(breaks=seq(2006,2020,1))
```

In contrast to minor shoplifting incidents, the number of major shoplifting incidents in the **Northeast Avalon** region is clearly **increasing** over time.

## Motor vehicle thefts

For the purposes of studying motor vehicle thefts, only the theft of the vehicle, not the vehicle's cost, will be considered.

```{r}
  data |>
  ggplot(aes(reorder(geography,total_theft_of_motor_vehicle),total_theft_of_motor_vehicle)) +
  geom_col(aes(fill="Thefts",color="Thefts")) + 
  labs(
    title="Total Motor Vehicle Thefts by Region",
    x="Region",
    y="Thefts",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  coord_flip()
```

The **Northeast Avalon** region experienced the largest total number of motor vehicle thefts. What about the general trend over time?

```{r}
  data |>
    filter(geography == "Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_theft_of_motor_vehicle)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Motor Vehicle Thefts",
      subtitle="Northeast Avalon",
      x="Year",
      y="Thefts",
      caption="Data from Open Data Newfoundland and Labrador",
    )
```

It appears that in the **Northeast Avalon** region, the number of total motor vehicle thefts is trending **downward** over time.

## Thefts from motor vehicles

How do the regions rank in terms of thefts from motor vehicles? In the case of thefts under $5,000:

```{r}
  data |>
  ggplot(aes(reorder(geography,theft_5_000_or_under_from_a_motor_vehicle),theft_5_000_or_under_from_a_motor_vehicle)) +
  geom_col(aes(fill="Thefts",color="Thefts")) + 
  labs(
    title="Total Motor Vehicle Thefts Under $5,000",
    x="Region",
    y="Thefts",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  coord_flip()
```
In the case of thefts above $5,000:

```{r}
  data |>
  ggplot(aes(reorder(geography,theft_over_5_000_from_a_motor_vehicle),theft_over_5_000_from_a_motor_vehicle)) +
  geom_col(aes(fill="Thefts",color="Thefts")) + 
  labs(
    title="Total Motor Vehicle Thefts Over $5,000",
    x="Region",
    y="Thefts",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  coord_flip()
```

In both cases, the **Northeast Avalon** region leads in terms of thefts from motor vehicles by a wide margin.

# Firearms and Weapons Offences

## How do the total number of firearms offences vary by region? 

```{r}
  data |>
    ggplot(aes(reorder(geography,total_firearms_use_of_discharge_pointing),total_firearms_use_of_discharge_pointing)) +
    geom_col(aes(color="Violations",fill="Violations")) +
    labs(
      title="Total Firearms Violations by Region",
      x = "Region",
      y = "Violations",
    ) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    coord_flip()
```

The **Northeast Avalon** region vastly outstrips the other regions in terms of firearms offences. Again, the question arises as to whether or not this is part of a trend for the region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_firearms_use_of_discharge_pointing)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Firearms Violations",
      subtitle="Northeast Avalon",
      x="Year",
      y="Violations",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    scale_x_continuous(breaks=seq(2006,2020,1))
```

There is a clear, **positive** trend in the number of firearms offences in the Northeast Avalon region. 

What about the use of weapons in general?

## How does the count of total weapons violations vary by region (firearms included)?

```{r}
  data |>
    ggplot(aes(reorder(geography,total_weapons_violations),total_weapons_violations)) +
    geom_col(aes(color="Violations",fill="Violations")) +
    labs(
      title="Total Weapons Violations by Region",
      x = "Region",
      y = "Violations",
    ) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    coord_flip()
```
Once again, we see that the **Northeast Avalon** ranks first, this time in terms of total weapons violations.

## Is the large number of total weapons violations in the Northeast Avalon part of a trend?

```{r}
  data |>
    filter(geography == "Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,total_weapons_violations)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Weapons Violations",
      subtitle="Northeast Avalon",
      x="Year",
      y="Violations",
      caption="Data from Open Data Newfoundland and Labrador",
    ) + 
    scale_x_continuous(breaks=seq(2006,2020,1))
```

It is uncertain whether or not the high numbers of total weapons violations in the region amount to a trend.

# Criminal Harassment and Threats

## How do the total number of incidents of criminal harassment vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,criminal_harassment))) +
  geom_col(aes(y=criminal_harassment,fill="Criminal Harassment",color="Criminal Harassment")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Incidents of Criminal Harassment by Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region leads in terms of incidences of criminal harassment. 

## How do the total number of incidents of uttering threats vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,uttering_threats))) +
  geom_col(aes(y=uttering_threats,fill="Threats",color="Threats")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Incidents of Uttering Threats by Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region dominates in terms of incidents of uttering threats.

## How do the total number of threatening phone calls break down by region?

```{r}
  title <- "Total Incidents of Threatening or Harassing Phone Calls By Region" |>
  str_wrap(60)

  data |>
  ggplot(aes(x=reorder(geography,threatening_or_harassing_phone_calls))) +
  geom_col(aes(y=threatening_or_harassing_phone_calls,fill="Threats",color="Threats")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title=title,
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```
Once again, the **Northeast Avalon** dominates in terms of total incidents, this time with respect to threatening or harassing phone calls.

## How have the total incidents of criminal harassment varied over time in the Northeast Avalon region?

```{r}
  data |>
    filter(geography == "Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,criminal_harassment)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    labs(
      title="Total Incidents of Criminal Harassment",
      subtitle="Northeast Avalon",
      x="Year",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) + 
    scale_x_continuous(breaks=seq(2006,2020,1))
```

Clearly, there has been a **positive** trend in the number of threatening or harassing phone calls over time, with respect to the **Northeast Avalon** region.

# Fraud and Identity Theft

## How do the regions vary in terms of total counts of fraud?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,fraud))) +
  geom_col(aes(y=fraud,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Incidents of Fraud by Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region dominates in terms of total counts of fraud by a wide margin.

## How do incidents of identity theft vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,identity_theft))) +
  geom_col(aes(y=identity_theft,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Incidents of Identity Theft by Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

Identity theft is clearly a prevalent issue in the **Northeast Avalon** region. It is often claimed that this form of crime is on the rise; it would be interesting to see if this is truly the case.

## Is there a trend in the total counts of identity theft in the Northeast Avalon region?

The metadata for the data set used in this study indicate that identity fraud and theft are both new violation codes as of January 2010. For the purposes of studying counts of identity theft, then, only years greater than or equal to 2010 will be considered.

```{r}
  data |>
    filter(geography == "Northeast Avalon, R.N.C.") |>
    filter(year >= 2010) |>
    ggplot(aes(year,identity_theft)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Identity Theft",
      subtitle="Northeast Avalon",
      x = "Year",
      y = "Incidents"
    )
```

It does indeed appear that incidents of identity theft are **increasing** over time, at least in the case of the **Northeast Avalon** region.

## How do the incidents of identity fraud vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,identity_fraud))) +
  geom_col(aes(y=identity_fraud,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Incidents of Identity Fraud by Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

In this case, the **Gander-Lewisporte** district is host to the largest number of incidents of identity fraud. 

## Are the high numbers of incidents part of a greater trend? 

Note that identity fraud was made a new violation code in January 2010. As such, we will only consider years greater than or equal to 2010 for the purposes of this study.

```{r}
  data |>
    filter(geography == "Gander-Lewisporte District, RCMP") |>
    filter(year >= 2010) |>
    ggplot(aes(year,identity_fraud)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Identity Fraud",
      subtitle="Gander-Lewisporte District",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

## How do the charges of counterfeiting vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,counterfeiting))) +
  geom_col(aes(y=counterfeiting,fill="Charges",color="Charges")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Charges of Counterfeiting by Region",
    x="Region",
    y="Charges",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The most charges for counterfeiting were laid in the **Northeast Avalon** region.

# Property Crime

## How do the total incidents of property crime vary by region?

```{r}
  data |>
  ggplot(aes(x=reorder(geography,total_property_crime_violations))) +
  geom_col(aes(y=total_property_crime_violations,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Property Crime Violations by Region",
    x="Region",
    y="Violations",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region leads in terms of total property crime violations.

## Is there a trend in the number of total property crime violations in the Northeast Avalon region?

```{r}
  data |>
    filter(geography == "Northeast Avalon, R.N.C.") |>
    ggplot(aes(year,identity_fraud)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Property Crime Violations",
      subtitle="Northeast Avalon",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There does appear to be a **positive** trend in the number of total property crime violations in the **Northeast Avalon** region.

## How do the various forms of property crime rank in the Northeast Avalon region?

```{r}
  data |>
    filter(geography=="Northeast Avalon, R.N.C.") |>
    ggplot(aes(year)) +
    geom_line(aes(y=total_breaking_and_entering,color="Breaking and Entering")) +
    geom_line(aes(y=possession_of_stolen_property,color="Possession of Stolen Property")) +
    geom_line(aes(y=arson,color="Arson")) +
    geom_line(aes(y=trespass_at_night,color="Tresspass at Night")) +
    geom_line(aes(y=proceeds_of_crime,color="Proceeds of Crime")) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  scale_y_log10() +
  theme(legend.position="bottom") +
  labs(
    title="Property Crimes by Type",
    subtitle="Northeast Avalon",
    x="Year",
    y="Charges",
    caption="Data from Open Data Newfoundland and Labrador",
    color="Legend",
  ) +
  guides(color=guide_legend(nrow=2, byrow=TRUE)) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

In the case of the **Northeast Avalon** region, breaking and entering is by far the most common property crime committed.

# Summary

In almost **every** category of crime studied, whether it be robberies, firearms offences or others, the dominance of the **Northeast Avalon** region in terms of sheer numbers cannot be ignored. This is perhaps an expected result, however; the **Northeast Avalon** is home to over [200,000](https://www150.statcan.gc.ca/n1/daily-quotidien/200131/t002c-eng.htm) people, nearly half of the entire population of the province. The remaining regions each comprise a considerably smaller percentage of the population; as such, the prevalence of the **Northeast Avalon** on the preceding plots may simply be a result of the region's high population. 

The fact remains that the **Northeast Avalon** region dominates in terms of sheer numbers, when it comes to a broad spectrum and variety of crimes committed. In almost every case, the region shows an upward trend in the type of crime under study.

This study has obvious weaknesses:

1. The trends in crime for the **Northeast Avalon** were investigated due to the sheer number and volume of crimes committed in the region. They may or may not be representative of the trends for the other regions, or the province as a whole.

2. This study only considers charges laid with respect to incidents and violations. It does **not** take into consideration the following court proceedings and their outcomes (e.g. whether or not a person is found innocent or guilty of what they are charged of).

# Conclusion

The **Northeast Avalon** region of Newfoundland and Labrador, being the most populated in the province, is also host to the majority of its crime. Over a 14 year period and a broad spectrum, crime has shown to be increasing in the region over time with few exceptions. This study demonstrates that in terms of sheer numbers, the **Northeast Avalon** region dominates in terms of a wide variety of crimes; not only that, but the amount of those crimes tend to be increasing. This has profound implications for policing in the province; police forces face the challenge of a rapidly changing "landscape" of crime in Newfoundland and Labrador.

