---
title: "Crime in Newfoundland and Labrador (2006-2020)"
output:
  html_document:
    df_print: paged
---

```{r}
  library(tidyverse)
  library(janitor)
  library(ggthemes)
  library(scales)

  # Disable the display of scientific notation.
  options(scipen=999)

  data <- read.csv("https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=17253")

  data <- data|>
    janitor::clean_names() |>
    mutate(geography=factor(geography)) |>
    # Focus upon actual incidents of crime.
    filter(data_type=="actual_incidents") |>
    # We will restrict our area of study to the police regions of Newfoundland
    # and Labrador.
    filter(
      geography != "Yukon",
      geography != "Saskatchewan",
      geography != "Quebec",
      geography != "Prince Edward Island",
      geography != "Ontario",
      geography != "Nunavut",
      geography != "Northwest Territories",
      geography != "New Brunswick",
      geography != "Manitoba",
      geography != "Canada",
      geography != "British Columbia",
      geography != "Alberta",
      geography != "Nova Scotia",
      geography != "Newfoundland and Labrador",
    ) |>
    select(!c('geography_type','geography_note','data_type')) |>
    mutate(geography = str_replace(geography,"Northeast Avalon R.N.C","Northeast Avalon, R.N.C.")) |>
    pivot_longer(
    cols = c(assault_level_1,assault_level_2_with_weapon_or_causing_bodily_harm,assault_level_3_aggravated),
    names_to = "assault_type",
    values_to = "assault_number",
  ) |>
  mutate(assault_type = case_when(
    assault_type == "assault_level_1" ~ "Level 1",
    assault_type == "assault_level_2_with_weapon_or_causing_bodily_harm" ~ "Level 2 (With a Weapon or Causing Bodily Harm)",
    assault_type == "assault_level_3_aggravated" ~  "Level 3 (Aggravated)",
  )) |>
    pivot_longer(
    cols = c(sexual_assault_level_1,sexual_assault_level_2_with_a_weapon_or_bodily_harm,sexual_assault_level_3_aggravated),
    names_to = "sexual_assault_type",
    values_to = "sexual_assault_number",
  ) |>
  mutate(sexual_assault_type = case_when(
    sexual_assault_type == "sexual_assault_level_1" ~ "Level 1",
    sexual_assault_type == "sexual_assault_level_2_with_a_weapon_or_bodily_harm" ~ "Level 2 (With a Weapon or Causing Bodily Harm)",
    sexual_assault_type == "sexual_assault_level_3_aggravated" ~  "Level 3 (Aggravated)",
  )) |>
  pivot_longer(
    cols=c(theft_5_000_or_under_from_a_motor_vehicle,theft_over_5_000_from_a_motor_vehicle),
    names_to="motor_vehicle_theft_type",
    values_to="motor_vehicle_theft_amount",
  ) |>
  mutate(motor_vehicle_theft_type = case_when(
    motor_vehicle_theft_type == "theft_5_000_or_under_from_a_motor_vehicle" ~ "$5,000 or Under",
    motor_vehicle_theft_type == "theft_over_5_000_from_a_motor_vehicle" ~ "Over $5,000"
  )) |>
  pivot_longer(
      cols=c(impaired_operation_causing_bodily_harm,impaired_operation_causing_death),
      names_to = "impaired_operation_type",
      values_to = "impaired_operation_number",
    ) |>
    mutate(impaired_operation_type=case_when(
      impaired_operation_type == "impaired_operation_causing_bodily_harm" ~ "Causing Bodily Harm",
      impaired_operation_type == "impaired_operation_causing_death" ~ "Causing Death"
    ))  
```

This [RStudio](https://resources.rstudio.com/categories/rstudio-ide/) notebook studies crime in Newfoundland and Labrador over a 14 year period, using a data set made available by **Open Data Newfoundland and Labrador** on their website.

# Violent Crime

## Which region experienced the largest number of total violent criminal code violations?

We can get a sense for the scale of violent crime in the province by examining the total number of violent criminal code violations in each region.

```{r}
  data |>
    distinct(geography, .keep_all = TRUE) |>
    ggplot(aes(reorder(geography,total_violent_criminal_code_violations), total_violent_criminal_code_violations)) +
    geom_col(aes(fill="Violations",color="Violations")) +
    labs(
      title="Violent Criminal Code Violations By Region",
      x="Region",
      y="Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    theme(legend.position = "none") +
    scale_color_colorblind() +
    scale_fill_colorblind()
```

It would seem that the **Northeast Avalon** region leads in terms of the number of total violent criminal code violations.

This is the case for one region of the province, however. What about for the province as a whole?

## Is there a trend in the number of total violent criminal code violations for the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_violent_criminal_code_violations=sum(total_violent_criminal_code_violations)) |>
    ggplot(aes(year,total_violent_criminal_code_violations)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Violent Criminal Code Violations",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is an apparent, **upward** trend in the number of violent criminal code violations across the province. Let us now examine specific forms of violent crime and their numbers.

## Which region experienced the largest number of homicides?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,homicide), homicide)) +
    geom_col(aes(fill="Homicides",color="Homicides")) +
    labs(
      title="Total Homicides By Region",
      x="Region",
      y="Homicides",
      subtitle="(2006-2020)",
      caption="Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    theme(legend.position = "none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    scale_y_continuous(breaks=breaks_pretty(n=3))
```

The **Northeast Avalon** region experienced the largest number of homicides.

## Which region leads in assaults against peace officers?

```{r}
data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(reorder(geography,assault_peace_officer),assault_peace_officer)) +
  geom_col(aes(color="Assaults",fill="Assaults")) +
  coord_flip() +
  labs(
    title="Assaults Against Peace Officers By Region",
    x="Region",
    y="Assaults",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

Here we see that the **Northeast Avalon** region leads in terms of assaults against peace officers. This certainly fits with the general trend of increasing violent crime within the region. It is concerning, however, to note the number of incidents in which even officers of the law became the victims of violent crimes.

## What kind and number of assaults were committed against members of the public in the province?

Assaults are ranked on a scale. There are three levels:

- level 1,
- level 2 (involving a weapon or causing bodily harm),
- level 3 (aggravated).

```{r}
  data |>
  ggplot(aes(year)) +
  geom_col(aes(y=assault_number,fill=assault_type),position="dodge") +
  theme(
    legend.position="bottom"
  ) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(
    title="Total Assaults By Type",
    subtitle="Newfoundland and Labrador",
    x="Year",
    y="Assaults",
    fill="Legend",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

**Level 1** assaults account for more than any other level.

# Sex Crimes

## How do the regions vary in terms of sexual assaults?

Sexual assaults are "ranked" by three levels:

- level 1,
- level 2 (committed with a weapon or causing bodily harm),
- level 3 (aggravated).

```{r}
  data |>
    distinct(geography,year,sexual_assault_type,.keep_all=TRUE) |>
    filter(sexual_assault_type=="Level 1") |>
    group_by(year) |>
    
    ggplot(aes(reorder(geography,sexual_assault_number))) +
    geom_col(aes(y=sexual_assault_number,color="Assaults",fill="Assaults")) +
    coord_flip() +
    labs(
      title="Sexual Assaults (Level 1) By Region",
      x="Region",
      y="Assaults",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    theme(
      legend.position="none"
    ) +
    scale_color_colorblind() +
    scale_fill_colorblind()
```

The **Northeast Avalon** region ranks first in terms of total number of sexual assaults (level 1).

```{r}
  data |>
    distinct(geography,year,sexual_assault_type,.keep_all=TRUE) |>
    filter(sexual_assault_type=="Level 2 (With a Weapon or Causing Bodily Harm)") |>
    ggplot(aes(reorder(geography,sexual_assault_number))) +
    geom_col(aes(y=sexual_assault_number,color="Assaults",fill="Assaults")) +
    coord_flip() +
    labs(
      title="Sexual Assaults (Level 2) By Region",
      x="Region",
      y="Assaults",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    theme(
      legend.position="none"
    ) +
    scale_color_colorblind() +
    scale_fill_colorblind()
```

Again, the **Northeast Avalon** leads in terms of sexual assaults (level 2), this time involving a weapon or bodily harm.

```{r}
  data |>
    distinct(geography,year,sexual_assault_type,.keep_all=TRUE) |>
    filter(sexual_assault_type=="Level 3 (Aggravated)") |>
    ggplot(aes(reorder(geography,sexual_assault_number))) +
    geom_col(aes(y=sexual_assault_number,color="Assaults",fill="Assaults")) +
    coord_flip() +
    labs(
      title="Sexual Assaults (Level 3) By Region",
      x="Region",
      y="Assaults",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    theme(
      legend.position="none"
    ) +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    scale_y_continuous(breaks=breaks_pretty()) 
```

The **Labrador District** experienced the greatest number of aggravated sexual assaults (level 3).

## How do the levels of sexual assault vary across the province?

```{r}
  data |>
  ggplot(aes(year)) +
  geom_col(aes(y=sexual_assault_number,fill=sexual_assault_type),position="dodge") +
  theme(
    legend.position="bottom"
  ) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(
    title="Total Sexual Assaults By Type",
    subtitle="Newfoundland and Labrador",
    x="Year",
    y="Assaults",
    fill="Legend",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

The most common type of sexual assault across the province is **level 1.**

## How are children affected by sex crimes in Newfoundland and Labrador?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,total_sexual_violations_against_children),total_sexual_violations_against_children)) +
    geom_col(aes(fill="Violations",color="Violations")) +
    labs(
      title="Sexual Violations Against Children By Region",
      x = "Region",
      y = "Sexual Violations",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The **Northeast Avalon** district experienced the most total sexual violations against children.

## Is there a trend in sexual violations against children in the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_sexual_violations_against_children=sum(total_sexual_violations_against_children)) |>
    ggplot(aes(year,total_sexual_violations_against_children)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Sexual Violations Against Children",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is a clear, **positive** trend in the number of sexual violations against children.

## How do the charges of prostitution vary by region?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,total_prostitution),total_prostitution)) +
    geom_col(aes(fill="Charges",color="Charges")) +
    labs(
      title="Charges of Prostitution By Region",
      x = "Region",
      y = "Charges",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

In this case the **Northeast Avalon** region outnumbers all others in terms of charges laid for prostitution.

## How do the incidents of indecent acts vary by region?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,indecent_acts),indecent_acts)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    labs(
      title="Indecent Acts By Region",
      x = "Region",
      y = "Incidents",
      caption = "Data from Open Data Newfoundland and Labrador"
    ) +
    coord_flip() +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    theme(legend.position="none")
```

The majority of crimes involving indecent acts were committed in the **Northeast Avalon** region.

# Robbery and Theft

## How did the number of robberies rank across the different regions over time?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,total_robbery),total_robbery)) +
    geom_col(aes(fill="Robberies",color="Robberies")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Robberies By Region",
      x="Region",
      y="Robberies",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

We see in this case that the **Northeast Avalon** experienced the most robberies over time; in fact, the region outstrips the others by a wide margin in terms of this form of crime.

## Is there a trend in the number of robberies in the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_robbery=sum(total_robbery)) |>
    ggplot(aes(year,total_robbery)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Robberies",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Robberies",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There has been a clear, **positive** trend in the number of robberies over time.

## How did the various regions vary in terms of shoplifting incidents?

```{r}
  data |>
    distinct(geography,year,shoplifting_5_000_or_under,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,shoplifting_5_000_or_under),shoplifting_5_000_or_under)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Shoplifting Incidents (Under $5,000)",
      x="Region",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

Here again, we see that the **Northeast Avalon** dominates by a wide margin in terms of the number of total minor shoplifting incidents over time.

## What about incidents in which over $5,000 of property was stolen?

```{r}
  data |>
    distinct(geography,year,shoplifting_over_5_000,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,shoplifting_over_5_000),shoplifting_over_5_000)) +
    geom_col(aes(fill="Incidents",color="Incidents")) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    labs(
      title="Shoplifting Incidents (Over $5,000)",
      x="Region",
      y="Incidents",
      caption="Data from Open Data Newfoundland and Labrador",
    ) +
    coord_flip()
```

Again, we see that the **Northeast Avalon** region vastly outnumbers its neighbors in terms of shoplifting incidents, even when those incidents involve much more expensive property.

## Is the number of thefts from motor vehicles above $5,000 increasing or decreasing across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(shoplifting_over_5_000=sum(shoplifting_over_5_000)) |>
    ggplot(aes(year,shoplifting_over_5_000)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Shoplifting (Above $5,000)",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

The number of major shoplifting incidents in the province is clearly **increasing** over time.

## How do the number of motor vehicle thefts vary by region?

For the purposes of studying motor vehicle thefts, only the theft of the vehicle, not the vehicle's cost, will be considered.

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(reorder(geography,total_theft_of_motor_vehicle),total_theft_of_motor_vehicle)) +
  geom_col(aes(fill="Thefts",color="Thefts")) + 
  labs(
    title="Motor Vehicle Thefts By Region",
    x="Region",
    y="Thefts",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  coord_flip()
```

The **Northeast Avalon** region experienced the largest total number of motor vehicle thefts. 

## Is there a trend in total motor vehicle thefts across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_theft_of_motor_vehicle=sum(total_theft_of_motor_vehicle)) |>
    ggplot(aes(year,total_theft_of_motor_vehicle)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Motor Vehicle Thefts",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Thefts",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

The total number of motor vehicle thefts appears to be **decreasing** over time across the province.

## Is theft from motor vehicles more common below or above $5,000?

```{r}
  data |>
  ggplot(aes(year)) +
  geom_col(aes(fill=motor_vehicle_theft_type,y=motor_vehicle_theft_amount),position="dodge") +
  guides(
    guides(fill=guide_legend(nrow=3,byrow=TRUE))
  ) +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Motor Vehicle Thefts By Type",
    subtitle="Newfoundland and Labrador",
    x="Year",
    y="Thefts",
    fill="Legend",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  theme(
    legend.position="bottom"
  ) +
  scale_x_continuous(breaks=seq(2006,2020,1))
```

When it comes to thefts from motor vehicles, thefts below \$5,000 are far more common than those above \$5,000.

## What are the trends for theft from motor vehicles across the province?

For thefts from motor vehicles below $5,000:

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    filter(motor_vehicle_theft_type=="$5,000 or Under") |>
    summarize(motor_vehicle_theft_amount=sum(motor_vehicle_theft_amount)) |>
    ggplot(aes(year,motor_vehicle_theft_amount)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Theft from Motor Vehicles ($5,000 or Under)",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

Thefts from motor vehicles below \$5,000 are clearly **decreasing** with time. What about thefts above \$5,000?

```{r}
  data |>
    distinct(geography,year,motor_vehicle_theft_type,.keep_all=TRUE) |>
    group_by(year) |>
    filter(motor_vehicle_theft_type=="Over $5,000") |>
    summarize(motor_vehicle_theft_amount=sum(motor_vehicle_theft_amount)) |>
    ggplot(aes(year,motor_vehicle_theft_amount)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Theft from Motor Vehicles (Over $5,000)",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

The number of high-value thefts from motor vehicles in the province is clearly **increasing** with time.

# Firearms and Weapons Offences

## How do the total number of firearms offences vary by region? 

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,total_firearms_use_of_discharge_pointing),total_firearms_use_of_discharge_pointing)) +
    geom_col(aes(color="Violations",fill="Violations")) +
    labs(
      title="Firearms Violations By Region",
      x = "Region",
      y = "Violations",
    ) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    coord_flip()
```

The **Northeast Avalon** region vastly outstrips the other regions in terms of firearms offences. 

## Is there a trend in the total number of firearms offences for the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_firearms_use_of_discharge_pointing=sum(total_firearms_use_of_discharge_pointing)) |>
    ggplot(aes(year,total_firearms_use_of_discharge_pointing)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Firearms Violations",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is a clear, **positive** trend in the number of firearms offences. 

What about the weapons violations in general?

## How does the count of total weapons violations vary by region (firearms included)?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,total_weapons_violations),total_weapons_violations)) +
    geom_col(aes(color="Violations",fill="Violations")) +
    labs(
      title="Weapons Violations By Region",
      x = "Region",
      y = "Violations",
    ) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    coord_flip()
```

Once again, we see that the **Northeast Avalon** ranks first, this time in terms of total weapons violations.

## Is the large number of total weapons violations in the Northeast Avalon part of a trend across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_weapons_violations=sum(total_weapons_violations)) |>
    ggplot(aes(year,total_weapons_violations)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Weapons Violations",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

In this case, it is clear that the number of total weapons violations across the province is **rising** over time.

## How do the total number of charges of weapon possession vary by region?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    ggplot(aes(reorder(geography,possession_of_weapons),possession_of_weapons)) +
    geom_col(aes(color="Charges",fill="Charges")) +
    labs(
      title="Weapons Possession Charges By Region",
      x = "Region",
      y = "Charges",
    ) +
    theme(legend.position="none") +
    scale_color_colorblind() +
    scale_fill_colorblind() +
    coord_flip()
```

The **Northeast Avalon** region leads in terms of charges for weapons possession. 

## Is this part of a larger trend across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(possession_of_weapons=sum(possession_of_weapons)) |>
    ggplot(aes(year,possession_of_weapons)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Charges of Possession of Weapons",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Charges",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There does appear to be a **positive** trend in the number of total weapons possession charges.

# Criminal Harassment and Threats

## How do the total number of incidents of criminal harassment vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,criminal_harassment))) +
  geom_col(aes(y=criminal_harassment,fill="Criminal Harassment",color="Criminal Harassment")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Criminal Harassment By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region leads in terms of incidences of criminal harassment. 

## How do the total number of incidents of uttering threats vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,uttering_threats))) +
  geom_col(aes(y=uttering_threats,fill="Threats",color="Threats")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Uttering Threats By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region dominates in terms of incidents of uttering threats.

## How do the total number of threatening phone calls vary by region?

```{r}
  title <- "Total Incidents of Threatening or Harassing Phone Calls By Region" |>
  str_wrap(50)
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,threatening_or_harassing_phone_calls))) +
  geom_col(aes(y=threatening_or_harassing_phone_calls,fill="Threats",color="Threats")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title=title,
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

Once again, the **Northeast Avalon** dominates in terms of total incidents, this time with respect to harassing or threatening phone calls.

## How have the total incidents of criminal harassment varied over time across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(criminal_harassment=sum(criminal_harassment)) |>
    ggplot(aes(year,criminal_harassment)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Criminal Harassment",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

Incidences of criminal harassment have been **increasing** with time for the province as a whole.

# Extortion, Fraud and Identity Theft

## How do incidents of extortion vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,extortion))) +
  geom_col(aes(y=extortion,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Extortion By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

Once again, the **Northeast Avalon** region ranks first, this time in incidents of extortion.

## How do the regions vary in terms of total incidents of fraud?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,fraud))) +
  geom_col(aes(y=fraud,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Fraud By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region dominates in terms of total counts of fraud by a wide margin.

## How do incidents of identity theft vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,identity_theft))) +
  geom_col(aes(y=identity_theft,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Identity Theft By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

Identity theft is clearly a prevalent issue in the **Northeast Avalon** region.

## Is there a trend in the amount of identity theft across the province?

Identity theft and fraud were introduced as violation codes in January of 2010, according to the [metadata](https://opendata.gov.nl.ca/public/opendata/filedownload/?file-id=41) accompanying the data set used in this study. As such, only years after 2010 (inclusive) will be considered.

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    filter(year >= 2010) |>
    summarize(identity_theft=sum(identity_theft)) |>
    ggplot(aes(year,identity_theft)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Identity Theft",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

The plot demonstrates a clear, **positive** trend in the number of incidents of identity theft across the province.

## How do the incidents of identity fraud vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,identity_fraud))) +
  geom_col(aes(y=identity_fraud,fill="Incidents",color="Incidents")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Identity Fraud By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

In this case, the **Gander-Lewisporte** district is host to the largest number of incidents of identity fraud. 

## Is there a trend in identity fraud with respect to the province as a whole?

Again, since identity fraud as a violation code was only introduced in 2010, in this case only years after 2010 (inclusive) will be considered.

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    filter(year >= 2010) |>
    summarize(identity_fraud=sum(identity_fraud)) |>
    ggplot(aes(year,identity_fraud)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Identity Fraud",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There appears to be a clear, **positive** trend in the number of incidents of identity fraud across the province.

## How do the charges of counterfeiting vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,counterfeiting))) +
  geom_col(aes(y=counterfeiting,fill="Charges",color="Charges")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Charges of Counterfeiting By Region",
    x="Region",
    y="Charges",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The most charges for counterfeiting were laid in the **Northeast Avalon** region.

## What is the trend for charges of counterfeiting across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    filter(year >= 2010) |>
    summarize(identity_fraud=sum(identity_fraud)) |>
    ggplot(aes(year,identity_fraud)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Identity Fraud",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

# Property Crime

## How do the total incidents of property crime vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,total_property_crime_violations))) +
  geom_col(aes(y=total_property_crime_violations,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Property Crime Violations By Region",
    x="Region",
    y="Violations",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region leads in terms of total property crime violations.

## Is there a trend in the number of total property crime violations in the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_property_crime_violations=sum(total_property_crime_violations)) |>
    ggplot(aes(year,total_property_crime_violations)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Property Crime Violations",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Violations",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is a definite, **negative** trend in the number of property crime violations in the province.

## How do the number of breaking and entering charges vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,total_breaking_and_entering))) +
  geom_col(aes(y=total_breaking_and_entering,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Breaking and Entering Charges By Region",
    x="Region",
    y="Charges",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region vastly outstrips its neighbors in terms of breaking and entering charges laid.

## Is there a trend in the number of breaking and entering charges for the province as a whole?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(total_breaking_and_entering=sum(total_breaking_and_entering)) |>
    ggplot(aes(year,total_breaking_and_entering)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Breaking and Entering Charges",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Charges",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

In this case, there is a clear, **downward** trend in the number of breaking and entering charges laid across the province.

## How do incidents of arson vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,arson))) +
  geom_col(aes(y=arson,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Incidents of Arson By Region",
    x="Region",
    y="Incidents",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

We see that the **Northeast Avalon** region leads in total incidents of arson.

## Is there a trend in the number of arson incidents across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(arson=sum(arson)) |>
    ggplot(aes(year,arson)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Incidents of Arson",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Incidents",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

The number of incidents of arson is clearly trending **downward** across the province.

# On the Road

## How do the incidents of impaired driving vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,total_impaired_driving))) +
  geom_col(aes(y=total_impaired_driving,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Impaired Driving Charges By Region",
    x="Region",
    y="Charges",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** leads in terms of total impaired driving charges.

## What is the trend in total impaired driving charge across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all = TRUE) |>
    group_by(year) |>
    summarize(total_impaired_driving=sum(total_impaired_driving)) |>
    ggplot(aes(year,total_impaired_driving)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Impaired Driving Charges",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Charges",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is a definite **increase** in the number of impaired driving charges over time.

## How do the number of impaired driving charges causing injury or death vary over time?

```{r}
  data |>
    distinct(geography,year,impaired_operation_type,.keep_all=TRUE) |>
    ggplot(aes(year)) +
    geom_col(aes(y=impaired_operation_number,color=impaired_operation_type,fill=impaired_operation_type),position="dodge") +
  theme(
    legend.position="bottom"
  ) +
  scale_x_continuous(breaks=seq(2006,2020,1)) +
  labs(
    title="Total Impaired Driving Incidents By Type",
    x="Year",
    y="Incidents",
    fill="Legend",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  guides(
    color="none"
  ) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

Thankfully, the number of impaired driving incidents in Newfoundland and Labrador that result in injury or death remain relatively low, compared to the number of impaired drivers on the road.

## How does the number of prohibited drivers on the road vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,driving_while_prohibited))) +
  geom_col(aes(y=driving_while_prohibited,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Prohibited Driver Charges By Region",
    x="Region",
    y="Drivers",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** leads in terms of drivers who are arrested while driving, when they are already prohibited from doing so.

## Is there a trend in the number of prohibited drivers across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |> 
    summarize(driving_while_prohibited=sum(driving_while_prohibited)) |>
    ggplot(aes(year,driving_while_prohibited)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Driving While Prohibited Charges",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Charges",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

It would appear that the number of prohibited drivers on the road is **increasing** with time.

# Drugs and Controlled Substances

## How do the arrests for cannabis possession vary across regions?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,posession_cannabis))) +
  geom_col(aes(y=posession_cannabis,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Arrests for Cannabis Possession By Region",
    x="Region",
    y="Arrests",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** is the leader in terms of arrests for cannabis possession.

## Is there a trend in the number of arrests for cannabis possession across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(posession_cannabis=sum(posession_cannabis)) |>
    ggplot(aes(year,posession_cannabis)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Arrests for Cannabis Possession",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Arrests",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There has been a general trend **downward** in the number of arrests for cannabis possession. In particular there was a steep drop in the number of arrests in 2018; this may coincide with the [legalization of cannabis in Canada](https://en.wikipedia.org/wiki/Legal_history_of_cannabis_in_Canada) in October of 2018.

## How do the number of arrests for the posession of cocaine vary by region?

```{r}
  data |>
  distinct(geography,year,.keep_all=TRUE) |>
  ggplot(aes(x=reorder(geography,possession_cocaine))) +
  geom_col(aes(y=possession_cocaine,fill="Violations",color="Violations")) +
  theme(legend.position="none") +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  labs(
    title="Total Arrests for Cocaine Possession By Region",
    x="Region",
    y="Arrests",
    caption="Data from Open Data Newfoundland and Labrador",
  ) +
  coord_flip()
```

The **Northeast Avalon** region ranks first in terms of arrests for cocaine possession.

## Is there a trend in the number of arrests for cocaine possession across the province?

```{r}
  data |>
    distinct(geography,year,.keep_all=TRUE) |>
    group_by(year) |>
    summarize(possession_cocaine=sum(possession_cocaine)) |>
    ggplot(aes(year,possession_cocaine)) +
    geom_line() +
    geom_smooth(se=FALSE,method="lm") +
    scale_x_continuous(breaks=seq(2006,2020,1)) +
    labs(
      title="Total Arrests for Cocaine Possession",
      subtitle="Newfoundland and Labrador",
      x = "Year",
      y = "Arrests",
      caption="Data from Open Data Newfoundland and Labrador"
    )
```

There is a clear, upward trend in the number of arrests for cocaine possession over time in the province.

# Summary

The dominance of the **Northeast Avalon** region in terms of sheer numbers of crimes committed cannot be ignored. This result is perhaps to be expected, however; the **Northeast Avalon** is home to over [200,000](https://www150.statcan.gc.ca/n1/daily-quotidien/200131/t002c-eng.htm) people, nearly half of the entire population of the province. This may, in part, explain why the region dominates in terms of total crimes committed across a broad spectrum of offences. 

This study has obvious weaknesses:

1. This study only considers charges laid with respect to incidents and violations. It does **not** take into consideration the following court proceedings and their outcomes (e.g. whether or not a person is found innocent or guilty of what they are charged of).

# Conclusion

The **Northeast Avalon** region of Newfoundland and Labrador, being the most populated in the province, is also host to the majority of its crime. However, this study has demonstrated that the problem of growing crime rates is not isolated to that region alone. The number of offences across the province is **increasing** for a broad spectrum of criminal activity. This will continue to pose a challenge for law enforcement.